### 2016-08-08 AnalysisNote.tex
Initial commit

\documentclass[a4paper,12pt,twoside]{article}
        \usepackage{graphicx,type1cm,eso-pic,color}
        \usepackage{hyperref}
        \usepackage{longtable}
        \usepackage{amsmath,amssymb} 
        \usepackage[latin1]{inputenc}
        \usepackage[OT1]{fontenc}
        \usepackage{setspace}
        \usepackage{epigraph}

\hypersetup{
     backref=true,    %permet d'ajouter des liens dans...
     pagebackref=true,%...les bibliographies
     hyperindex=true, %...les index.
     colorlinks=true, %colorise les liens
     breaklinks=true, %permet le retour à la ligne dans les liens trop longs
     urlcolor= blue, %couleur des hyperliens
     linkcolor= black, %couleur des liens internes
     bookmarks=true, %créé des signets pour Acrobat
     bookmarksopen=true,            %si les signets Acrobat sont créés,
                                    %les afficher complètement.
     pdftitle={Hadronization Analysis Note (EG2)}, %informations apparaissant dans
     pdfauthor={Raphael Dupre},     %dans les informations du document
     pdfsubject={Physics}          %sous Acrobat.
}

\setlength{\hoffset}{-18pt}  	
\setlength{\oddsidemargin}{20pt} 	% Marge gauche sur pages impaires
\setlength{\evensidemargin}{20pt} 	% Marge gauche sur pages paires
\setlength{\marginparwidth}{54pt} 	% Largeur de note dans la marge
\setlength{\textwidth}{440pt} 		% Largeur de la zone de texte (17cm)
\setlength{\voffset}{18pt}	        % Bon pour DOS
\setlength{\marginparsep}{7pt} 	 	% Séparation de la marge
\setlength{\topmargin}{0pt} 	        % Pas de marge en haut
\setlength{\headheight}{13pt} 		% Haut de page
\setlength{\headsep}{0pt}              % Entre le haut de page et le texte
\setlength{\footskip}{27pt} 	        % Bas de page + séparation
\setlength{\textheight}{600pt} 	 	% Hauteur de la zone de texte (25cm)

\newcommand{\xbp}{$x_{Bj}$}
\newcommand{\xb}{$x_{Bj}~$}
\newcommand{\ptp}{$p_\perp^2$}
\newcommand{\pt}{$p_\perp^2~$}
\newcommand{\dptp}{$\Delta \langle p_\perp^2 \rangle$}
\newcommand{\dpt}{$\Delta \langle p_\perp^2 \rangle ~$}

\makeatother

\title{Study of the hadronization of $\pi^-$}

\author{Rapha\"el \textsc{Dupr\'e}, 
Lamiaa \textsc{El Fassi}\thanks{also {\it Rutgers University}} , 
Kawtar \textsc{Hafidi} ... \\ 
{\it Physics Division, Argonne National Laboratory,} \\
{\it Argonne, Illinois 60439, USA} }

\date{\today}

\begin{document}

\maketitle

%doublespacing
%onehalfspacing
%\linenumbers
\renewcommand{\baselinestretch}{1.10}

\begin{abstract}
Hadronization is the process by which energetic quarks transform into 
colorless hadrons. The process is non-perturbative, therefore, only 
qualitative understanding can be achieved, and using phenomenological models 
is necessary. It is believed that hadronization has two distinct phases: 
the quark propagation phase, where the quark emit gluons until it neutralizes 
its color by forming a pre-hadron and the hadron formation phase, where the 
pre-hadron evolves into a hadron. Those process are respectively characterized 
by the production time and the formation time. By comparing hadron production 
on different nuclei, one can measure variables with different sensitivities 
to the hadronization phases, the two principal variables are: the transverse 
momentum broadening, that is believed to be tightly connected to quark energy 
loss and the multiplicity ratio, which is a measure of the hadron attenuation 
in the medium.

The goals of the experiment E-02-104 are to study both quark propagation
and hadronization dynamic using various nuclear target.
In this analysis we focus on two particles, the negative pion and the proton.
Our goals are to constrain the theoritical models and extract the production
time using precise multidimensional analysis of $\pi^+$ and $\pi^-$ to enlight possible
new physical processes with a precise analysis of proton production in nuclei.

\end{abstract}

\newpage

\tableofcontents

\newpage

\include{intro}

\include{chap5}

\include{chap6}

\bibliography{hadro}
\bibliographystyle{plain}

\end{document}