### 2016-10-20 AnalysisNote.tex
Changes from Lamiaa

\documentclass[a4paper,12pt,twoside]{article}
        \usepackage{graphicx,type1cm,eso-pic,color}
        \usepackage{hyperref}
        \usepackage{longtable}
        \usepackage{amsmath,amssymb} 
        \usepackage[latin1]{inputenc}
        \usepackage[OT1]{fontenc}
        \usepackage{setspace}
        \usepackage{epigraph}

\hypersetup{
     backref=true,    %permet d'ajouter des liens dans...
     pagebackref=true,%...les bibliographies
     hyperindex=true, %...les index.
     colorlinks=true, %colorise les liens
     breaklinks=true, %permet le retour à la ligne dans les liens trop longs
     urlcolor= blue, %couleur des hyperliens
     linkcolor= black, %couleur des liens internes
     bookmarks=true, %créé des signets pour Acrobat
     bookmarksopen=true,            %si les signets Acrobat sont créés,
                                    %les afficher complètement.
     pdftitle={Hadronization Analysis Note (EG2)}, %informations apparaissant dans
     pdfauthor={Raphael Dupre et al.},     %dans les informations du document
     pdfsubject={Nuclear Physics}          %sous Acrobat.
}

\setlength{\hoffset}{-18pt}  	
\setlength{\oddsidemargin}{20pt} 	% Marge gauche sur pages impaires
\setlength{\evensidemargin}{20pt} 	% Marge gauche sur pages paires
\setlength{\marginparwidth}{54pt} 	% Largeur de note dans la marge
\setlength{\textwidth}{440pt} 		% Largeur de la zone de texte (17cm)
\setlength{\voffset}{18pt}	        % Bon pour DOS
\setlength{\marginparsep}{7pt} 	 	% Séparation de la marge
\setlength{\topmargin}{0pt} 	        % Pas de marge en haut
\setlength{\headheight}{13pt} 		% Haut de page
\setlength{\headsep}{0pt}              % Entre le haut de page et le texte
\setlength{\footskip}{27pt} 	        % Bas de page + séparation
\setlength{\textheight}{600pt} 	 	% Hauteur de la zone de texte (25cm)

\newcommand{\xbp}{$x_{Bj}$}
\newcommand{\xb}{$x_{Bj}~$}
\newcommand{\ptp}{$p_\perp^2$}
\newcommand{\pt}{$p_\perp^2~$}
\newcommand{\dptp}{$\Delta \langle p_\perp^2 \rangle$}
\newcommand{\dpt}{$\Delta \langle p_\perp^2 \rangle ~$}

\makeatother

\title{Study of the hadronization of charged pions}

\author{
  Rapha\"el \textsc{Dupr\'e} \\ 
  {\it Institut de Physique Nucl\'eaire, Orsay, France} \\
  Lamiaa \textsc{El Fassi} \\ 
  {\it Mississippi State University, Mississippi State,} \\ 
  {\it Mississippi 39762, USA} \\
  Kawtar \textsc{Hafidi} \\ 
  {\it Physics Division, Argonne National Laboratory,} \\ 
  {\it Argonne, Illinois 60439, USA} \\
  William \textsc{Brooks} \\
  {\it Universidad T\'ecnica Federico Santa Mar\'ia,} \\ 
  {\it Valpara\'iso, Chile} 
}

\date{\today}

\begin{document}

\maketitle

%doublespacing
%onehalfspacing
%\linenumbers
\renewcommand{\baselinestretch}{1.10}

\begin{abstract}
Hadronization is the process by which energetic quarks transform into colorless hadrons. The process is non-perturbative, therefore, only a qualitative understanding based on phenomenological models can be achieved. By comparing hadron production on different nuclei, one can measure the most sensitive variables to the hadronization phases, the transverse momentum broadening, that is believed to be tightly connected to the quark energy loss, and the multiplicity ratio, which is a measure of the hadron attenuation in the nuclear medium.

The goal of the experiment E-02-104 was the study of the quark propagation and hadronization dynamics using various nuclear targets. In this note, we focus on the production of two charged particles, positive and negative pions, to constrain theoretical models using a precise multidimensional analysis.

\end{abstract}

\newpage

\tableofcontents

\newpage

\include{intro}

\include{chap5}

\include{chap6}

\bibliography{hadro}
\bibliographystyle{ieeetr}

\end{document}